---
title: "Reproducible Research Assignment 1"
author: "Tamás Jakab"
date: "November 20, 2016"
output: word_document
---
Downloading, unzipping and loading the csv file into RStudio:

```{r}
link <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

temp <- tempfile()
download.file(link, temp)
data <- read.csv(unz(temp, "activity.csv"),header = T, sep = ",")
unlink(temp)
```

Calculating the total number of steps per day
```{r}
perday <- aggregate(.~ date, data = data, sum)
```

Histogram of the total number of steps taken each day
```{r}
hist(perday$steps, breaks = 10, main = "Total number of steps each day", xlab = "Steps")
```

Mean and median number of steps taken each day
```{r}
mean(perday$steps)
median(perday$steps)
```

Calculating the average steps per interval
```{r}
mean.inter <- aggregate(.~ interval, data = data, mean)
```

Time series plot of the average number of steps taken
```{r}
plot(x=mean.inter$interval, y =mean.inter$steps, type = "l", main = "Time series plot of the average number of steps taken", xlab = "Intervals", ylab = "Average steps")
```

5-minute interval containing the maximum number of steps
```{r}
mean.inter[mean.inter$steps == max(mean.inter$steps),1]
```

Calculating the total number of rows with NAs
```{r}
sum(is.na(data))
```

Imputing missing values
```{r}
library(plyr)
data2 <- data
impute <- function(x, fun) {
      missing <- is.na(x)
      replace(x, missing, fun(x[!missing]))
}
data2 <-ddply(data2, ~ interval, transform, steps = impute(steps, mean))
```

Calculating the total number of steps per day(no NAs)
```{r}
perday2 <- aggregate(.~ date, data = data2, sum)
```

Histogram of the total number of steps taken each day(no NAs)
```{r}
hist(perday2$steps, breaks = 10, main = "Total number of steps each day", xlab = "Steps")
```

Mean and median number of steps taken each day(no NAs)
```{r}
mean(perday2$steps)
median(perday2$steps)
```

Creating new factor variable
```{r}
data2$w <- weekdays(as.Date(data2$date))
data2$weekday <- ifelse(data2$w == "Saturday" | data2$w == "Sunday", data2$w <-"weekend",data2$w <-"weekday")
data2$w <- NULL
```

Creating average steps based on weekdays/weekend and ploting them
```{r}
par(mfrow = c(2,1))
mean.interval.weekday <- aggregate(steps~ interval, data = subset(data2, weekday == "weekday"), mean)
plot(x=mean.interval.weekday$interval, y =mean.interval.weekday$steps, type = "l", main = "Weekday", xlab = "Intervals", ylab = "Average steps")

mean.interval.weekend <- aggregate(steps~ interval, data = subset(data2, weekday == "weekend"), mean)
plot(x=mean.interval.weekend$interval, y =mean.interval.weekend$steps, type = "l", main = "Weekend", xlab = "Intervals", ylab = "Average steps")
```
